---
title: "iteration_and_listcols"
author: "Iris Hart"
date: "2024-10-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


load packages 
```{r}
library(tidyverse)
library(rvest)
```

```{r}
set.seed(1)
```

### Here are some lists 
```{r}
l = list(
  vec_numeric = 1:4,
  unif_sample = runif(100),
  mat = matrix(1:8, nrow = 2, ncol = 4, byrow = T),
  summary = summary(rnorm(100))
)

l

l$mat

l[["mat"]][1, 3]

l[[1]]
l[[4]]

```

make a list that is hopefully a bit more useful 

```{r}
list_norm = 
  list(
    a = rnorm(20, 0, 5),
    b = rnorm(20, 4, 5),
    c = rnorm(20, 0, 10),
    d = rnorm(20, 4, 10)
  )

list_norm[["b"]]
```

lets reuse the function we used last time 
```{r}
mean_and_sd = function(x) {
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  out_df = tibble(
    mean = mean_x,
    sd = sd_x
  )
  
  return(out_df)
}

```

Lets use the function to take mean and standard deviation 
```{r}
mean_and_sd(list_norm[["a"]])
mean_and_sd(list_norm[["b"]])
mean_and_sd(list_norm[["c"]])
mean_and_sd(list_norm[["d"]])
```

## use a for loop 

create ouput list, and run a for loop 
```{r}
output = vector("list", length = 4)

for(i in 1:4) {
  output[[i]] = mean_and_sd(list_norm[[i]])
}

output
```

## Do the same thing 
but with 'map' instead
```{r}
output = map(list_norm, mean_and_sd)
```

```{r}
output = 
  output = map(list_norm,mean_and_sd) |>
  bind_rows()

output = map(list_norm, mean_and_sd)

output = map_dbl(list_norm, IQR)
```

## LIST COLUMNS !!

```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norm
  )

listcol_df

listcol_df |>
  filter(name %in% c("a", "b"))

listcol_df |> 
  select(-samp)

```

```{r}
listcol_df[["samp"]][["a"]]
```


compute mean and sd 
```{r}
mean_and_sd(listcol_df[["samp"]][["a"]])
mean_and_sd(listcol_df[["samp"]][["b"]])

map(listcol_df[["samp"]], mean_and_sd)
```

ADD A LIST COLUMN

```{r}
listcol_df |>
  mutate(
    output = map(samp, mean_and_sd),
    iqr = map_dbl(samp, IQR))

listcol_df |>
  mutate(
    output = map(samp, mean_and_sd),
    iqr = map_dbl(samp, IQR)) |>
  select(-samp) |> 
    unnest(output)
  
```


### NSDUHS
```{r}
nsduh_table_format =function(html, table_num, table_name) {
  
  out_table = 
    html |>
    html_table() |>
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value"))
 
 return(out_table)

}
```

We need to import teh html, and then extract the correct variables
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)
```

```{r}
nsduh_table_format(html = nsduh_html, table_num = 1)
nsduh_table_format(html = nsduh_html, table_num = 4)
nsduh_table_format(html = nsduh_html, table_num = 5)
```

```{r}
nsduh_df = 
  tibble(
    drug = c("marj", "cocaine", "heroin"),
    table_n = c(1, 4, 5)
  ) |> 
  mutate(table = map(table_n, nsduh_table_format, html = nsduh_html)) |>
    unnest(table)

nsduh_df |> 
  filter(State == "New York")
```




